---
title: "lab_1_in_class"
author: "Laura Wolf"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:
- basic review (data wrangling)
- intro to spatial data in R with sf
- intro to shiny apps

1. Attach packages

```{r load_packages}

library(tidyverse)
library(sf)
library(leaflet)
library(tmap)

```

2. Review: reading, exploring, wrangling and visualizing data

a. Read in the data
```{r load_data}

wine <- read_csv("wine_data.csv")

```

b. Wrangling 
```{r wrangling}

wine_new <- wine %>% #"and then". never overwrite original data
  select(country, province, winery, region_1, points, price) %>% #select=columns. can be used for ordering
  rename(state = province) %>% # = sets something equal. == looks for a match
  filter(state == "California" | state == "Oregon" | state == "Washington") %>%  #filter is for rows. | means "or". selects only the states california, oregon, and washington.
  mutate(ppd = points/price) %>% #mutate can change a column or add a new one. adds a column for points per dollar
  arrange(-ppd) #arrange stores values in decreasing ppd order

```

c. Use group_by() and summarize() to find summary statistics by group
```{r summary_stats}

wine_summary <- wine_new %>% 
  group_by(state) %>% #create groups that I won't see, but I know exist to group data by states. used to calculate things based on groups. can have multiple levels of groupings.
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE), #create a data table that shows mean ppd in the groups we created
    max_ppd = max(ppd, na.rm = TRUE) #adds max ppd to table
  )
  

```

d. A few gg-graphics
```{r histograms}

wine_hists <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) + #need the aes to know to specify color by state. fill = bar color
  facet_wrap(~state, scales = "free") #puts each on its own plot, default is to have the same scale

wine_hists

```

new graph: scatterplot graph of price vs. points, with color determined by state, adding a linear model trendline on top of it

```{r scatterplot}

wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  geom_point(aes(color = state), alpha = 0.5) + #points use "color" to set color of points
  geom_smooth(color = "black", method = "lm", se = FALSE) + #adds black trendline without standard error
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange")) +
  facet_wrap(~state) + #puts each on its own plot
  theme_light() #changes the display theme

wine_points

```

2. California counties (spatial data) using sf

```{r load_spatial_data}

ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016")

ca_land <- ca_counties %>% 
  select(NAME, ALAND) #keeps geometry!!! "sticky geometry"

ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% #load ca_pop_inc data
  rename(NAME = COUNTY) #renames the county column to "NAME" to match the column "NAME" in ca_land. values in the cells match.

ca_df <- full_join(ca_land, ca_pop_inc) %>% #safest best because it keeps everything
  select(NAME, MedFamilyIncome)

#plot(ca_counties)

```

d. Use geom_sf to make a map in ggplot

```{r map}

ca_income <- ggplot(ca_df) +
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>% 
  addTiles() %>% 
  addPolygons()

```

